Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-06-16 WSSSPE2_report/WSSSPE2_review.tex
Use en-dash for number range

The workshop featured two keynote addresses. In the opening keynote
presentation, Kaitlin Thaney of the Mozilla Science Lab talked about her
organization's work and policy to enable and support sustainable and
reproducible scientific research through the open web. The second keynote
speaker was Neil Chue Hong of Software Sustainability Institute. He shined a
light on how scientific software is prevalently driving advances in many science
and engineering fields. Both keynote speeches spawned further discussion among
workshop participants on the crucial notion \emph{sustainability} in the theme
of our workshop.

@@

Finally, after agreeing that these complementary training models were needed, the group discussed sustainability of
training, and how the leading groups in this space are teaching only a small fraction of the community that
needs and wants training.  Most graduate programs in the sciences do not currently incorporate these approaches
in their graduate curricula, although there are an increasing number of quantitatively oriented courses around
analysis and modeling.  These still, however, generally omit engineering practices such as version control, unit
and regression testing, and software modularity and abstraction, often because the instructors themselves in the
domain sciences are not familiar with these techniques.  Thus, students who are being trained in these approaches
are doing so through short 2-3 day training workshops such as Software and Data Carpentry, rather than through
semester-long graduate education courses at their University, which tend to focus on statistical and modeling
techniques.  Hybrid programs like the three week Open Science for Synthesis (OSS) training that combine the three
(science, quantitative techniques, and software engineering) into an holistic course serve part of the need but
reach only a few researchers at this time.  Thus, the group concluded that significant strides needed to be made
in coordinating these trainings so that they are complementary, on increasing their size and scope, and on
better integrating them directly into graduate education programs.

>>>>

The workshop featured two keynote addresses. In the opening keynote
presentation, Kaitlin Thaney of the Mozilla Science Lab talked about her
organization's work and policy to enable and support sustainable and
reproducible scientific research through the open web. The second keynote
speaker was Neil Chue Hong of Software Sustainability Institute. He shed
light on how scientific software is prevalently driving advances in many science
and engineering fields. Both keynote speeches spawned further discussion among
workshop participants on the crucial notion of \emph{software sustainability}
in the theme of our workshop.

@@

Finally, after agreeing that these complementary training models were needed, the group discussed sustainability of
training, and how the leading groups in this space are teaching only a small fraction of the community that
needs and wants training.  Most graduate programs in the sciences do not currently incorporate these approaches
in their graduate curricula, although there are an increasing number of quantitatively oriented courses around
analysis and modeling.  These still, however, generally omit engineering practices such as version control, unit
and regression testing, and software modularity and abstraction, often because the instructors themselves in the
domain sciences are not familiar with these techniques.  Thus, students who are being trained in these approaches
are doing so through short 2--3 day training workshops such as Software and Data Carpentry, rather than through
semester-long graduate education courses at their University, which tend to focus on statistical and modeling
techniques.  Hybrid programs like the three week Open Science for Synthesis (OSS) training that combine the three
(science, quantitative techniques, and software engineering) into an holistic course serve part of the need but
reach only a few researchers at this time.  Thus, the group concluded that significant strides needed to be made
in coordinating these trainings so that they are complementary, on increasing their size and scope, and on
better integrating them directly into graduate education programs.
